---
title: "2nd assignment"
output:
  word_document: default
date: "2022-10-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("plyr")
#install.packages("FSA")
#install.packages("FSAdata")
#install.packages("magrittr")
#install.packages("dplyr")
#install.packages("plotrix")
#install.packages("ggplot2")
#install.packages("moments")

library(plyr)
library(FSA)
library(FSAdata)
library(magrittr)
library(dplyr)
library(plotrix)
library(ggplot2)
library(moments)
```


## Q1 Print your name at the top of the script. Include the prefix: “Plotting Basics:” 
```{r}
print("Plotting Basics: Udaikiran")
```

## Q2 Import libraries including: plyr, FSA, FSAdata, magrittr,  dplyr, plotrix, ggplot2, and moments 

```{r}
library(plyr)
library(FSA)
library(FSAdata)
library(magrittr)
library(dplyr)
library(plotrix)
library(ggplot2)
library(moments)
```

## Q3 Load the BullTroutRML2 dataset (https://www.rdocumentation.org/packages/FSAdata/versions/0.3.9/topics/BullTroutRML2)

```{r}
data("BullTroutRML2")
```

## Q4 Print the first and last 3 records from the dataset 

```{r}
head(BullTroutRML2,n=3)
tail(BullTroutRML2,n=3)
```

## Q5 Filter out all records except those from Harrison Lake 

```{r}
BullTroutRML2 %<>% dplyr::filter(lake=="Harrison")
print(BullTroutRML2)
```

## Q6 Display the first and last 3 records from the filtered dataset 

```{r}
headtail(BullTroutRML2,n=3)
```


## Q7 Display the structure of the filtered dataset 
```{r}
str(BullTroutRML2)
```

## Q8 Display the summary of the filtered dataset and save it as <t> 

```{r}
summary(BullTroutRML2)
t<-summary(BullTroutRML2)
```

## Q9 Create a scatterplot for “age” (y variable) and “fl” (x variable) with the following 
##  specifications: 
##    • Limit of x axis is (0,500) 
##    • Limit of y axis is (0,15) 
##    • Title of graph is “Plot 1: Harrison Lake Trout 
##    • X axis label is “Age (yrs)” 
##    • Y axis label is “Fork Length (mm)” 
##    • Use a small filled circle for the plotted data points 

```{r}
x_q9<- BullTroutRML2$fl
y_q9<-BullTroutRML2$age
plot(x_q9,y_q9, ylim=c(0, 15), xlim=c(0,500), ylab="Age (yrs)", 
     xlab="Fork Length (mm)", pch=20, las=1, cex=.5, main="Plot 1: Harrison Lake Trout ")
```

## Q10 Plot an “Age” histogram with the following specifications  
##  specifications: 
##    • Y axis label is “Frequency”
##    • X axis label is “Age (yrs)” 
##    • Title of the histogram is “Plot 2: Harrison Fish Age Distribution” 
##    • The color of the frequency plots is “cadetblue”
##    • The color of the Title is “cadetblue” 

```{r}
hist(BullTroutRML2$age, ylim=c(0,15), xlim=c(0,15), ylab="Frequency", 
     xlab="Age (yrs)", main="Plot 2: Harrison Fish Age Distribution", col="cadetblue1")
```

## Q11 Create a overdense for “age” (y variable) and “fl” (x variable) with the following 
##  specifications: 
##    • Limit of x axis is (0,500) 
##    • Limit of y axis is (0,15) 
##    • Title of graph is “Plot 3: Harrison Density Shaded by Era" 
##    • X axis label is “Age (yrs)” 
##    • Y axis label is “Fork Length (mm)” 
##    • Use a small filled circle for the plotted data points 
##    • include two levels of shading for the “black” data points

```{r}
x_q11<- BullTroutRML2$fl
y_q11<-BullTroutRML2$age
plot(x_q11,y_q11, ylim=c(0, 15), xlim=c(0,500), ylab="Age (yrs)", 
     xlab="Fork Length (mm)", pch=20, col=rgb(0,0,0,1/2), las=1, cex=.5, main="Plot 3: Harrison Density Shaded by Era")
```



## Q12 Create a new object called “tmp” that includes the first 3 and last 3 records of the whole data set.

```{r}
tmp <- BullTroutRML2[c(1:3,31:33),]
tmp
```

## Q13  Display the “era” column in the new “tmp” object 

```{r}
tmp$era
```

## Q14 Create a pchs vector with the argument values for + and x. Then create a cols vector with the two elements “red” and “gray60” 

```{r}
pchs <- c(3,4)

cols <-c("red","gray60")
```

## Q15 Convert the tmp object values to numeric values. Then create a numeric numEra object from the tmp$era object 

```{r}
as.numeric(tmp$era)
numEra <- as.numeric(tmp$era)
```
## Q16 Associate the cols vector with the tmp era values 
```{r}
cols[tmp$era]
```


## Q17  Create a plot of “Age (yrs)” (y variable) versus “Fork Length (mm)” (x variable) with the 
##following specifications: 
##    • Limit of x axis is (0,500) 
##    • Limit of y axis is (0,15) 
##    • Title of graph is “Plot 4: Symbol & Color by Era” 
##    • X axis label is “Age (yrs)” 
##    • Y axis label is “Fork Length (mm)” 
##    • Set pch equal to pchs era values 
##    • Set col equal to cols era values 

```{r}
plot(age~fl,data=BullTroutRML2, ylim=c(0, 15), xlim=c(0,500), ylab="Age (yrs", 
     xlab="Fork Length (mm)", pch=pchs[era], col=cols[tmp$era], main="Plot 4: Symbol & Color By Era")
```
## Q18 Plot a regression line of the previous plot with a dashed line with width 2 and color “cadetblue” 

```{r}
cols[numEra]
plot(age~fl,data=BullTroutRML2, ylim=c(0, 15), xlim=c(0,500), ylab="Age (yrs", 
     xlab="Fork Length (mm)", pch=pchs[era], col=cols[numEra], main="Plot4C: Symbol & Color By Era")
regLine <-lm(age~fl, data=BullTroutRML2)
abline(regLine, col="cadetblue", lty=2, lwd=2)
```

## Q19 Place a legend of levels by era with pchs symbols in the top left of the plot with the following ##specifications: 
##    • Inset of 0.05 
##    • No box around the legend 
##    • Font size: 75% of nominal 

```{r}
plot(age~fl,data=BullTroutRML2, ylim=c(0, 15), xlim=c(0,500), ylab="Age (yrs", 
     xlab="Fork Length (mm)", pch=pchs[era], col=cols[numEra], main="Plot4C: Symbol & Color By Era")
regLine <-lm(age~fl, data=BullTroutRML2)
abline(regLine, col="cadetblue", lty=2, lwd=2)
legend("topleft", inset=0.05, legend=levels(BullTroutRML2$era),
       pch=pchs, col=cols, bty="n", cex=0.75)

```


